---
title: "Group 8 Project"
subtitle: "PIXE analysis on Maya blue in Prehispanic and colonial mural paintings"
author: "Awet, Connor, Lachlan, Pablo,"
format: html
editor: visual
---

```{r}
#| include: false
library("readxl")
Table1 <- read_excel("../Data/Table1_clean.xlsx")
Table1
Table1_raw <- read_excel("../Data/Table1_raw.xlsx")
Table1_raw
```

### Introduction

Back Ground on article

Aim of this analysis

### Method

Explain Data: Number of observations, variables, variables measured, and mention/reference table 1 from paper.

Statistical Methods we are going to use} (Can put code in here for what we do to show the method of doing it if needed but I will talk to lecturer about if we need to or not.

We are doing this on the elements, comparing them to each other to find patterns, etc. names down the rows of the table are just the observations

Methods we are doing based off lectures, can change if needed

1.  Descriptive Statistics (mean, median, sd, min, max of each variable)

```{r}

```

2.  Exploratory pairs plot 17x17 for the 17 variables

    1.  Histograms with notched box plots for each variable to show distribution

    2.  contours to show 3rd box plots (can also just do heat map for with contours painted on)

    3.  Mahalanobis-coloured scatter plots

        {All of this in one plot, as seen in the one in inference - 03.pdf

```{r}

```

3.  Covariance Matrix and Correlation Heat map

```{r}

```

```{r}

```

4.  Check normality

    1.  Mahalanobis Distance Plot

```{r}

```

```         
2.  Chi-square Q-Q plot
```

```{r}

```

5.  Principle component analysis (For the reproducing question)

    1.  Scree plot
    2.  scores plot
    3.  loadings

```{r}

```

6.  Cluster analysis

```{r}

```

### Results

1.  Show and explain through the analysis of what we described in the method. (Plots will go in here with explanations and thoughts under each.

<!-- -->

1.  Descriptive Statistics (mean, median, sd, min, max of each variable)

```{r}

```

2.  Exploratory pairs plot 17x17 for the 17 variables

```{r}

```

3.  Covariance Matrix and Correlation Heat map

```{r}

```

4.  Normality Check

```{r}

```

5.  Principle component analysis (For the reproducing question)

```{r}

```

6.  Cluster analysis

```{r}

```

### Discussion

Answering the questions from the document and any other conclusions we have from the analysis

1.  Can you find groups in these data?
2.  Can you reproduce Fig. 2? Do you agree with the conclusions the authors draw about the “zones”?

#### References {style="text-align: center;"}

> Andonie, R., and I. Dzitac. 2010. “How to Write a Good Paper in Computer Science and How Will It Be Measured by ISI Web of Knowledge.” International Journal of Computers, Communications & Control V (4): 432–44. https://doi.org/10.15837/ijccc.2010.4.2493.
>
> Frery, A. C., L. Gomez, and A. C. Medeiros. 2020. “A Badging System for Reproducibility and Replicability in Remote Sensing Research.” IEEE Journal of Selected Topics on Applied Earth Observations and Remote Sensing 13: 4988–95. https://doi.org/10.1109/JSTARS.2020.3019 418.
>
> Sánchez del Río, M., Martinetto, P., Solís, C. & Reyes-Valerio, C. (2006), ‘PIXE analysis on Maya blue in Prehispanic and colonial mural paintings’, Nuclear Instruments and Methods in Physics Research Section B: Beam Interactions with Materials and Atoms 249(1–2), 628–632.
